### 2015-03-20 tex-inputs/experiment.tex
Rough out experiment and results sections.

%!TEX root = ../paper.tex

\section{Experiment}

Recruited through journalist contacts.

Ran five experiments.

One-on-one with researcher.
Researcher limited interaction.
Demo cognitive walkthrough.
Cognitive walkthrough understanding task (Zimbabwe).

List of tasks.
Removed "Configure" from list of tasks after p2 because of confusion.

Live viewing by Tor Project developers.

Transcribed audio and made captioned screen recording videos.

\section{Results}

Common errors and confusion.

Apple Gatekeeper problems,
difficulty of fixing (concerns with the App Store).

Selected interesting quotes.

Number of new tickets created and fixed.